Package Interconnection

### 2021-10-10 paper/package_interconnection.Rmd
Add inline code to the text.

Interaction between packages could indicate of how developers utilize the functionality of one or more packages to create new features in their packages. According to @StanstrupJan2019TmTi, dependency and connectivity in terms of reusing codes and inheriting objects could reflect a well-connected and interacting community. In this paper, we observe the interaction limited to only between packages that are listed as DoE packages. Further, the type of interaction include the depended, imported, and suggested packages. Out of `r nrow(pkg_tidy)`, there are only `r nrow(pkg_nodes)` (`r round(nrow(pkg_nodes)/nrow(pkg_tidy)*100,2)`%) packages that interplay each other. This number is relatively lower compared to the percentage of interaction in the packages in Survey and Time Series Analysis CRAN Task View. We found that `r round(73/132*100,2)`% Survey packages that depends, imports, and suggests each other, while these interconnection is observed in `r round(212/338*100,2)`% Time Series Analysis packages.

>>>>

```{r}
# interaction in survey CRAN task view

list_pkg_survey <- ctv:::.get_pkgs_from_ctv_or_repos("OfficialStatistics", 
                                                repos = "http://cran.rstudio.com/")[[1]]
url <- "http://cran.rstudio.com/web/packages/packages.rds"
db <- readRDS(url(url)) %>% 
  as.data.frame()
survey_pkg <- db %>%
  filter(Package %in% list_pkg_survey) %>%
  select(Package, Depends, Imports, Suggests)

get_pkg_survey <- function(var){
  # select only the package name and depends/imports/suggests
  tmp_df <- survey_pkg %>%
    select(Package, var) %>%
    # separate every depends/imports/suggests
    cSplit(var, sep = " ,")

  df <- tmp_df %>%
    pivot_longer(c(2:ncol(tmp_df)), names_to = "type", values_to = "pkg") %>%
    filter(!is.na(pkg)) %>%
    # check if the depends/imports/suggests is also a DoE package (is in list_pkg)
    mutate(survey_pkg_related = ifelse(pkg %in% list_pkg_survey, "yes", "no")) %>%
    # filter only package whose depends/imports/suggests is a DoE
    filter(survey_pkg_related == "yes") %>%
    # tidy the data to only have the type: depends/imports/suggests
    cSplit("type", sep = "_") %>%
    select(-type_2) %>%
    rename(type = type_1)

survey_depends <- get_pkg_survey("Depends")
survey_imports <- get_pkg_survey("Imports")
survey_suggests <- get_pkg_survey("Suggests")

pkg_edges_survey <- rbind(survey_depends, survey_imports, survey_suggests)

survey_from <- pkg_edges_survey %>%
  distinct(Package)

# extract unique package that being depends/imports/suggests
survey_to <- pkg_edges_survey %>%
  distinct(pkg)

# create nodes from the two previous dataframe
pkg_nodes_survey <- full_join(survey_from, survey_to, by = c("Package" = "pkg"))

# Interaction in Time Series Analysis CRAN task view

list_pkg_ts <- ctv:::.get_pkgs_from_ctv_or_repos("TimeSeries", 
                                                repos = "http://cran.rstudio.com/")[[1]]

ts_pkg <- db %>%
  filter(Package %in% list_pkg_ts) %>%
  select(Package, Depends, Imports, Suggests)

get_pkg_ts <- function(var){
  # select only the package name and depends/imports/suggests
  tmp_df <- ts_pkg %>%
    select(Package, var) %>%
    # separate every depends/imports/suggests
    cSplit(var, sep = " ,")

  df <- tmp_df %>%
    pivot_longer(c(2:ncol(tmp_df)), names_to = "type", values_to = "pkg") %>%
    filter(!is.na(pkg)) %>%
    # check if the depends/imports/suggests is also a DoE package (is in list_pkg)
    mutate(survey_pkg_related = ifelse(pkg %in% list_pkg_ts, "yes", "no")) %>%
    # filter only package whose depends/imports/suggests is a DoE
    filter(survey_pkg_related == "yes") %>%
    # tidy the data to only have the type: depends/imports/suggests
    cSplit("type", sep = "_") %>%
    select(-type_2) %>%
    rename(type = type_1)

ts_depends <- get_pkg_ts("Depends")
ts_imports <- get_pkg_ts("Imports")
ts_suggests <- get_pkg_ts("Suggests")

pkg_edges_ts <- rbind(ts_depends, ts_imports, ts_suggests)

ts_from <- pkg_edges_ts %>%
  distinct(Package)

# extract unique package that being depends/imports/suggests
ts_to <- pkg_edges_ts %>%
  distinct(pkg)

# create nodes from the two previous dataframe
pkg_nodes_ts <- full_join(ts_from, ts_to, by = c("Package" = "pkg"))
```

Interaction between packages could indicate of how developers utilize the functionality of one or more packages to create new features in their packages. According to @StanstrupJan2019TmTi, dependency and connectivity in terms of reusing codes and inheriting objects could reflect a well-connected and interacting community. In this paper, we observe the interaction limited to only between packages that are listed as DoE packages. Further, the type of interaction include the depended, imported, and suggested packages. Out of `r nrow(pkg_tidy)`, there are only `r nrow(pkg_nodes)` (`r round(nrow(pkg_nodes)/nrow(pkg_tidy)*100,2)`%) packages that interplay each other. This number is relatively lower compared to the percentage of interaction in the packages in Survey and Time Series Analysis CRAN Task View. We found that `r round(nrow(pkg_nodes_survey)/nrow(survey_pkg)*100,2)`% Survey packages that depends, imports, and suggests each other, while these interconnection is observed in `r round(nrow(nrow(pkg_nodes_ts)/nrow(ts_pkg)*100,2)`% Time Series Analysis packages.
